---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

1.  Install two packages `gganimate` and `gifski` then restart Rstudio. Using the `Adult Census Income` data, make an animation using `geom_point` and `transition_states`.

```{r}
library(gganimate)
library(gifski)
library(ggplot2)
library(tidyverse)
library(knitr)
library(lubridate)
library(gapminder)
library(VIM)
```

```{r}
df <- read_csv("C:\\Users\\student\\Documents\\M461-Statistical-Analysis-with-R\\adult_census.csv")
```

```{r}
df %>% ggplot(
  aes (
    x = education.num, 
    y = , fnlwgt,
    color = sex)) + 
  geom_point()+
  transition_states(sex)+
  labs(title = 'Sex: {closest_state}')
```

2.  Using the `Adult Census Income` data, make an animation using `geom_bar` and `transition_states`.

```{r}
df %>% ggplot(
  aes(
    x = income,
    fill = race)) +
  geom_bar(position = 'fill') + 
  transition_states(sex) +
  labs(title = 'Sex: {closest_state}')
```

2.  Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19 in 2021.

```{r}
df <- read_csv("C:\\Users\\student\\Documents\\M461-Statistical-Analysis-with-R\\WHO-COVID-19-global-data.csv")
```

```{r}
class(df$Date_reported)
```

```{r}
df$Date_reported <- as.Date(df$Date_reported, format = "%m/%d/%Y")
```

```{r}
class(df$Date_reported)
```

```{r}
df$month <- month(df$Date_reported)
d1 <- df %>% 
  group_by(month, Country) %>% 
  summarise(mean = mean(New_deaths), .groups = 'drop')
d2 <- d1 %>% 
  group_by(month) %>% 
  mutate (rank = rank(-mean))
d3 <- d2 %>% 
  filter(rank <= 10)
```

```{r}
a1 <- d3 %>% 
  ggplot(aes(x = rank, y = mean, group = Country, fill = Country, label = Country)) + 
  geom_col() +
  geom_text(aes(
    y = mean,
    label = Country),
    hjust = 1.4) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_x_reverse() +
  labs(title = 'Month {closest_state}',
       x = '',
       y = 'Total Number of Deaths',
       fill = 'Country') +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  transition_states(month) +
  ease_aes("cubic-in-out")

animate(a1, nframes = 400)
  
```

2.  Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (<https://covidtracking.com/data/download/all-states-history.csv>) but you should make a different bar racev from ones in the slides

```{r}
df <- read_csv("C:\\Users\\student\\Documents\\M461-Statistical-Analysis-with-R\\Salary_data.csv")
```

```{r}
d1 <- df %>% 
  group_by(education, gender) %>% 
  summarise(mean = mean(loyalty), .groups = 'drop')
d2 <- d1 %>% 
  group_by(education) %>% 
  mutate (rank = rank(-mean))
d3 <- d2 %>% 
  filter(rank <= 10)
```

```{r}
a1 <- d3 %>% 
  ggplot(aes(x = rank, y = mean, group = gender, fill = gender, label = gender)) + 
  geom_col() +
  geom_text(aes(
    y = mean,
    label = gender),
    hjust = 1.4) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_x_reverse() +
  labs(title = 'education level {closest_state}',
       x = '',
       y = 'loyalty level',
       fill = 'gender') +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  transition_states(education) +
  ease_aes("cubic-in-out")

animate(a1, nframes = 400)
```
