---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 11: Predictive Modeling - Universal Framework"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

------------------------------------------------------------------------

# 1. Install the package mlbench and use the follows to import the data

```{r}
library(mlbench)
library(tidyverse)
library(caret)
data(PimaIndiansDiabetes)
df <- tibble(PimaIndiansDiabetes)
```

## 1.1 Train and Test a decision tree and a random forest with caret. Plot the variable importance of these models.

### 1.1.1 Decision Tree with Caret

```{r}
df <- df %>% rename(target = diabetes)

df <- df %>% 
  mutate(target = as.factor(target))
```

```{r}
splitIndex <- createDataPartition(df$target, p = 0.7,
                                  list = FALSE)
df_train <- df[splitIndex,]
df_test <- df[-splitIndex,]
```

```{r}
model1 <- train(target ~., data = df_train,
                method = "rpart2",
                maxdepth = 3)

pred1 <- predict(model1, df_test)

cm1 <- confusionMatrix(data = pred1, reference = df_test$target, positive = "pos")

cm1$overall[1]
```

```{r}
plot(varImp(model1))
```

### 1.1.2 Random Forest with Caret

```{r}
model2 <- train(target ~., data = df_train,
                method = "rf",
                ntree  = 1000)

pred2 <- predict(model2, df_test)

cm2 <- confusionMatrix(data = pred2, reference = df_test$target, positive = "pos")

cm2$overall[1]
```

```{r}
plot(varImp(model2))
```

# 2. Train and Test a `glmnet` model. Plot the variable importance of the model.

```{r}
model3 <- train(target ~., data = df_train,
                method = "glmnet")

pred3 <- predict(model3, df_test)

cm3 <- confusionMatrix(data = pred3, reference = df_test$target, positive = "pos")

cm3$overall[1]
```

**Accuracy = 76.96%**

```{r}
plot(varImp(model3))

```

# 3. Train and test two models that are not `ranger`, `glmnet` or `rpart` with caret. If possible, plot the variable importance of the model.

## 3.1 Support Vector Machines (SVM)

```{r}
library(caret)
library(e1071)

model4 <- train(target ~., data = df_train,
                method = "svmRadial")

pred4 <- predict(model4, df_test)

cm4 <- confusionMatrix(data = pred4, reference = df_test$target, positive = "pos")

cm4$overall[1]
```

Cannot plot variable importance

## 3.2 K-Nearest Neighbors(kNN)

```{r}
model5 <- train(target~., data = df_train,
                method = "knn",
                tuneGrid = data.frame(k = 1:10))

pred5 <- predict(model5, df_test)

cm5 <- confusionMatrix(data = pred5, reference = df_test$target, positive = "pos")

cm5$overall[1]
```

cannot plot variable importance
