---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Data Wrangging with `dplyr`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so.

```{r}
library(tidyverse)
```

------------------------------------------------------------------------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://covidtracking.com/data/download/all-states-history.csv). Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv.

```{r}
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```

------------------------------------------------------------------------

#### 3. Fix the date and ceate some new variables

`lubridate` is a package of the `tidyverse` packages. We will make uses of `lubridate` in this question.

-   Use the below codes to create `month`, `weekday` and `monthday` variables

```{r}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

------------------------------------------------------------------------

#### 4. Create new variables with `case_when`.

The function `case_when` is a good option to create a new variable from existing variable. For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable. `deathIncrease` is the number of daily new death by Covid19. The new variable `daily_death` takes three values: low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14). Please notice that this can also be done in a different way as shown in Assignment 3.

```{r}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  df$deathIncrease > 14 ~ 'high',
  TRUE ~ 'high'
)
```

-   Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month \> 20).

```{r}
df$month2 <- case_when(
  df$month <=10 ~ 'early_month',
  df$month <=20 ~ 'mid_month',
  df$month >20 ~ 'end_month',
  TRUE~'end_month'
)
```

-   Create variable `weekend` that takes two values: 1 if it's Saturday or Sunday or 0 otherwise.

```{r}
df$weekend <- case_when(
  df$weekday == 6 ~ '1',
  df$weekday == 7 ~ '1',
  TRUE ~ '0'
)
```

------------------------------------------------------------------------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.

```{r}
df %>%  select(-totalTestsViral)
```

------------------------------------------------------------------------

#### 6. Pipe Operator ( %\>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable. Pipe works very well with all the `tidyverse` packages. Refer to [these slides](../slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
sqrt(x)

sum(sqrt(x))

log(sum(sqrt(x)))

# log base 2 of 16
log(16, 2)

```

```{r}
x <- c(1:10)

x %>% sqrt %>% sum %>% log
```

------------------------------------------------------------------------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

-   calculate the average number of cases by dataqualitygrade

```{r}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) by `month`

```{r}
df %>% 
  group_by(month) %>% 
  summarise(median(positiveIncrease))
```

-   Calculate the average number of cases (`positiveIncrease`) by `month2`

```{r}
df %>% group_by(month2) %>% 
  summarise(mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) by `weekend`

```{r}
df %>% group_by(weekend) %>% 
  summarise(median(positiveIncrease))
```

------------------------------------------------------------------------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020.

```{r}
df$year <- year(df$date)

df %>% 
  filter(year == 2020) %>% 
  filter(month == 10 | month ==11) %>%
  group_by(weekend) %>% 
  summarise(positive_increase = median(positiveIncrease))
  
```

-   Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020

```{r}
df %>% 
  filter(year == 2020) %>% 
  filter(month == 9| month == 10| month ==11) %>% 
  group_by(month2) %>%
    summarise(positive_increase = mean(positiveIncrease))
```

-   Compare the average number of hospitalizations between weekdays and weekends in Summer 2020

```{r}
df %>% 
  filter(year == 2020) %>%
  filter(month == 6 | month ==7| month ==8) %>% 
  group_by(weekend) %>%
  summarise(hospitalizations = mean(hospitalizedIncrease))
```

-   Redo Questions 14 and 15 in [Assignment 3](assignment3.html) using the combos. Notice: you also need to use the data used in Assignment 3.

Find the median cases by weekdays in Rhode Island in 2021.

```{r}
df <- read_csv('https://bryantstats.github.io/math421/data/us-states.csv')
```

```{r}
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)

df$year <- year(df$date)
```

```{r}
df %>%
  filter(year == 2021) %>% 
  filter( state == 'Rhode Island') %>% 
  group_by(weekday) %>% 
  summarise(cases = median(cases))
```

Compare the median cases in Rhode Island in June, July, August, and September in 2021.

```{r}
df %>% 
  filter(year == 2021) %>% 
  filter(state == 'Rhode Island') %>%
  filter(month == 6 | month == 7| month == 8| month ==9) %>%
  group_by(month) %>%
  summarise(median_cases = median(cases))
```

------------------------------------------------------------------------

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.

```{r}
df %>%
  filter(weekday == 6|weekday ==7) %>% 
  group_by(month) %>% 
  summarise(death_count = sum(deaths)) %>%
  arrange(-death_count)
```

------------------------------------------------------------------------

#### 10. Use your own dataset and implement the follows functions or combos. You can use the [Adult Census Income](../data/adult_census.csv) or [Titanic data](../data/titanic.csv).

```{r}
df <- read_csv('https://bryantstats.github.io/math421/data/titanic.csv')
```

-   select

```{r}
#remove cabin

df %>%  select(-Cabin)
```

-   filter

```{r}
#filter by females who survived
df %>%  filter(Sex == 'female', Survived == 1)
```

-   mutate

```{r}
#crate a column that has a 1 if first class and a 0 if not
df <- df %>% mutate(first_class = ifelse(Pclass == 1, 1, 0) )
```

-   summarise

```{r}
#summarise median fare
df %>%  summarise (median_fare = median(Fare, na.rm = TRUE))
```

-   arrange

```{r}
#arrange by class
df %>% arrange(Pclass)
```

-   count

```{r}
#count by class
df %>% count(Pclass)
```

-   count + arrange

```{r}
#determine what class appears most
df %>% 
  count(Pclass) %>% 
  arrange(-n)
```

-   filter + count + arrange

```{r}
#determine what class appears most for females
df %>% 
  filter(Sex == 'female') %>% 
  count(Pclass) %>% 
  arrange(-n) 

```

-   group_by + summarise

```{r}
#mean fare by class
df %>% 
  group_by(Pclass) %>% 
  summarise(mean_fare = mean(Fare))
```

-   filter + group_by + summarise

```{r}
#mean age by class for women
df %>% 
  filter(Sex == 'female') %>% 
  group_by(Pclass) %>% 
  summarise(mean_age = mean(Age, na.rm = TRUE))
```

-   filter + group_by + summarise + arrange

```{r}
# median fare by class for men in descending order
df %>% 
  filter(Sex == 'male') %>% 
  group_by(Pclass) %>% 
  summarise(median_fare = median(Fare, na.rm = TRUE)) %>% 
  arrange(-median_fare)
```
