---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

------------------------------------------------------------------------

**Questions**

1.  **Use `read_csv` to import the data. Show the number of NAs for each columns.**

```{r}
library(tidyverse)
df <- read_csv('https://bryantstats.github.io/math421/data/adult_census_missing.csv')
```

```{r}
colSums(is.na(df))
```

2.  **Using function `aggr` the VIM package to plot the number of NAs for each column.**

```{r}
library(VIM)
aggr(df)
```

2.  **Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data.**

```{r}
table(df$age, useNA = "ifany")
```

*age has "NA"*

```{r}
table(df$workclass, useNA = "ifany")
```

*workclass has "?", "Not Known", "Unknown", "NA"*

```{r}
table(df$fnlwgt)
```

```{r}
table(df$education)
```

*education has "Not Known", "Unknown"*

```{r}
table(df$education.num)
```

```{r}
table(df$marital.status)
```

*"Not Known"*

```{r}
table(df$occupation)
```

*"??, "Not Known", "Unknown"*

```{r}
table(df$relationship)
```

*"Not Known", "Unknown"*

```{r}
table(df$race)
```

```{r}
table(df$sex)
```

*"Unknown"*

```{r}
table(df$capital.gain)
```

*"99999"*

```{r}
table(df$capital.loss)
```

```{r}
table(df$hours.per.week)
```

```{r}
table(df$native.country)
```

*"?", "Not Known", "Unknown"*

```{r}
table(df$income)
```

2.  **Replace all the forms of missing values found to NA**

```{r}
df <- replace (df, df == 'Not Known' |
                 df == 'Unknown' |
                 df == '?' |
                 df == 99999, NA)
```

2.  **Replot the number of NAs for each column.**

```{r}
colSums(is.na(df))
```

```{r}
library(VIM)
aggr(df)
```

2.  **Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs). How many rows left after removing?**

```{r}
df1 <- drop_na(df)
```

32,561 rows changed to 29,717 rows

2.  **Approach 2 to handle NAs: Fill all the NAs with the previous or next value. (Hint: Using `fill` function). Save the dataset after filling as a different data. The original data is unchanged (still have NAs).**

```{r}
df2 <- df
```

```{r}
df2 %>%  fill(age, workclass,education, marital.status, occupation, relationship, sex, capital.gain, native.country, .direction = 'updown')
```

2.  **Approach 3 to handle NAs: For numeric variable, replace the NAs by the median. For categorical variables, replace the NAs to the majority.**

```{r}
df3 <- df
```

```{r}
median_age <- median(df3$age, na.rm = TRUE)
df3$age <- replace_na(df3$age, median_age)
```

```{r}
majority_workclass <- names(which.max(table(df3$workclass)))
df3$workclass <- replace_na(df3$workclass, majority_workclass)
```

```{r}
majority_education <- names(which.max(table(df3$education)))
df3$education <- replace_na(df3$education, majority_education)
```

```{r}
majority_marital.status <- names(which.max(table(df3$marital.status)))
df3$marital.status <- replace_na(df3$marital.status, majority_marital.status)
```

```{r}
majority_occupation <- names(which.max(table(df3$occupation)))
df3$occupation <- replace_na(df3$occupation, majority_occupation)
```

```{r}
majority_relationship <- names(which.max(table(df3$relationship)))
df3$relationship <- replace_na(df3$relationship, majority_relationship)
```

```{r}
majority_sex <- names(which.max(table(df3$sex)))
df3$sex <- replace_na(df3$sex, majority_sex)
```

```{r}
median_capital.gain <- median(df3$capital.gain, na.rm = TRUE)
df3$capital.gain <- replace_na(df3$capital.gain, median_capital.gain)
```

```{r}
majority_native.country <- names(which.max(table(df3$native.country)))
df3$native.country <- replace_na(df3$native.country, majority_native.country)
```
