---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

## 1. Use the `PimaIndiansDiabetes` dataset. Use 15% data for testing. Use cross-validation with of 7 folds to tune random forest `(method='ranger')`. What are the parameters that produce the greatest accuracy? What is the testing accuracy.

### 1.1 load data set

```{r}
library(tidyverse)
library(caret)
library(mlbench)
data(PimaIndiansDiabetes)
df <- tibble(PimaIndiansDiabetes)
```

### 1.2 Use 15% data for testing

```{r}
# set target
df <- df %>% rename(target = diabetes)

# correct variables' types
df <- df %>% 
  mutate(target = as.factor(target))
```

```{r}
splitIndex <- createDataPartition(df$target, p = 0.85,
                                  list = FALSE)
df_train <- df[splitIndex,]
df_test <- df[-splitIndex,]
```

### 1.3 use cross validation with 7 folds to tune random forest

```{r}
tuneGrid <- expand.grid(
  mtry = 2:8,
  splitrule = c("gini", "extratrees"),
  min.node.size = c(1:10))

trControl = trainControl(method = "cv",
                         number = 7)

```

```{r}
forest_ranger <- train(target ~., data = df_train,
                      method = "ranger",
                      trControl = trControl,
                      tuneGrid = tuneGrid)
```

### 1.4 parameters & accuracy

```{r}
plot(forest_ranger)
```

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")

cm$overall[1]
```

## 2. Use the `PimaIndiansDiabetes` dataset. Go to <https://topepo.github.io/caret/available-models.html> and pick a classification model. Tune the classification model using cross-validation of 7 folds.

```{r}
tuneGrid <- expand.grid(
sigma = c(0.01, 0.05, 0.1), C = c(0.5, 1, 2))


trControl = trainControl(method = "cv",
                         number = 7)

svm_model <- train(target ~., data = df_train,
                      method = "svmRadial",
                      trControl = trControl,
                      tuneGrid = tuneGrid)

plot(svm_model)

pred <- predict(svm_model, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")

cm$overall[1]
```

## 3. Model Comparison) Use the `PimaIndiansDiabetes` dataset. Pick two models at [this link](https://topepo.github.io/caret/available-models.html) to compare using 7-fold cross validation method. Evaluate the accuracy of the final model on the test data.

```{r}

trControl = trainControl(method = "cv",
                         number = 7)

AdaBag <- train(target~., data = df_train,
                method = "AdaBag",
                trControl = trControl)

knn <- train(target~., data = df_train,
                method = "knn",
                trControl = trControl)

results <- resamples(list('Bagged AdaBoost' = AdaBag,
                          'K nearest neighbors' = knn))

bwplot(results)
```

```{r}
pred <- predict(AdaBag , df_test)

cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")

cm$overall[1]
```
