---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3.

------------------------------------------------------------------------

## Problems

```{r}
library(dplyr)
```

1.  **Create the following data frame**

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28, 34, 29, 42),
                Name = c("Tom", "Jack", "Steve", "Ricky"))
df
```

2.  **Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).**

```{r}
df <- read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")
```

3.  **Show the names of the variables in the data**

```{r}
names(df)
```

4.  **How many columns and rows the data have?**

```{r}
str(df)
```

61,942 rows & 5 columns

5.  **How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?**

```{r}
sum(is.na(df))
```

0 missing values

```{r}
colSums(is.na(df))
```

0 missing values by each column

no variable has missing values

6.  **What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.**

```{r}
class(df$date)
```

date is classified as character

```{r}
df$date = as.Date(df$date)
```

```{r}
class(df$date)
```

7.  **Capitalize the names of all the variables**

```{r}
names(df)[1] = "Date"
names(df)[2] = "State"
names(df)[3] = "Flips"
names(df)[4] = "Cases"
names(df)[5] = "Death"
```

8.  Find the average number of cases per day. Find the maximum cases a day.

```{r}
df_bydate <- df[, c('Date', 'Cases')]

```

```{r}
max(df$Cases)
```

9.  How many states are there in the data?

```{r}
n_distinct(df$State)
```

10. Create a new variable `weekdays` to store the weekday for each rows.

```{r}
df$Weekdays <- weekdays(df$Date)
```

11. Create the categorical variable `death2` variable taking the values as follows

`has_death` if there is a death that day

`no_death` if there is no death that day

```{r}
df$Death2 <- ifelse(df$Death==0,
                    'no_death',
                    'has_death')
```

Find the frequency and relative frequency of `no_death` and `has_death`.

```{r}
table(df$Death2)
```

```{r}
prop.table(table(df$Death2))
```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

```{r}
summary(df$Death)
```

11. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)
df$Death3 = case_when(df$Death < 1171 ~ 'low_death',
                    df$Death < 14912 ~ 'mid_death',
                    TRUE ~ 'high_death')
```

13. Find the average cases in Rhode Island in 2021

```{r}
df_RI_2021 <- df[(df$State == 'Rhode Island')&(df$Date >= '2021-01-01') & (df$Date <= '2021-12-31'),]

```

```{r}
mean(df_RI_2021$Cases)
```

14. Find the median cases by weekdays in Rhode Island in 2021

```{r}
df_RI_2021_weekday <- df_RI_2021[df_RI_2021$Weekdays == '']
```

13. Compare the median cases in Rhode Island in June, July, August and September in 2021.

14. Find your own dataset, import it and implement the following functions on the data

```{r}
df <- read.csv('https://bryantstats.github.io/math421/data/titanic.csv')
```

-   head

```{r}
head(df)
```

-   str

```{r}
str(df)
```

-   names

```{r}
names(df)
```

-   mean, min, max

```{r}
mean(df$Fare)
```

```{r}
min(df$Fare)
```

```{r}
max(df$Fare)
```

-   table

```{r}
table(df$Pclass)
```

-   is.na

```{r}
sum(is.na(df))
```

-   colSums

```{r}
colSums(is.na(df))
```

-   class

```{r}
class(df$Pclass)
```

-   cor

```{r}
cor(df$Pclass, df$Survived)
```

-   by

```{r}
by(df$Fare, df$Sex, mean)
```

-   ifelse

```{r}
df$Pclass2<- ifelse(df$Pclass ==1, 'First_Class',ifelse  (df$Pclass ==2, 'Second_Class', 'Third_Class'))
```

-   case_when

```{r}
df$Pclass3 <- case_when(df$Pclass2 == 'First_Class' ~ "First class", df$Pclass2 == 'Second_Class' ~ 'Second Class', TRUE ~ 'Third Class')
```

17. In the dataset in #16, practice the follows. You can reuse the code of 16.

-   Create a categorical variable from a continuous variable

based of Fare amount

```{r}
summary(df$Fare)
```

```{r}
df$Fare2 <- case_when(df$Fare < 8 ~ 'low_fare',
                      df$Fare < 31 ~ 'medium_fare',
                      TRUE ~ 'high_fare')
```

-   From one categorical variable, create a new categorical variable with fewer categories

based
