---
title: "Data Wrangling with Dplyr"
author: "Aileen Charron"
format: html
editor: visual
---

# Data Wrangling with Dplyr

## three ways to data wrangle

-   base r

-   dplyr (*most popular)*

-   data.table

## packages

-   tidyverse

-   ggplot2: visualization

-   dplyr: data wrangling

-   purrr: functional programming

## read_csv

use *read_csv* instead of *read.csv* to read the data

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(purrr)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')

```

-   *`read_csv`* is faster

-   *`read_csv`* recognizing the types of data better

```{r}
df <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')

str(df)
```

-   data prints out nicely with *`read_csv`*

```{r}
df
```

## Pipe (%\>%)

-   hotekey: ctrl + shift+ m

## pipe operator

-   is the same as f(x)

```{r}
x <- c(1:10)
```

conventional sum

```{r}
sum(x)
```

pipe

```{r}
x %>% sum
```

### nested function

conventional

```{r}
log(sum(x))
```

pipe

```{r}
x %>% sum %>% log
```

# dplyr

## five verbs of dplyr

-   *`select()`* selects columns from data

-   `filter()`filter rows of data

-   `summarise()` summarizes data (calculating summary statistics)

-   `arrange()` arrange data

-   `mutate()` creates new variables

### select

select: select columns

```{r}
df <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
df1 <-  select(df, state, date)
```

select with pipe

```{r}
df %>% select(state, date)
```

## five verbs of dplyr

-   select() selects columns from data

-   filter() filter rows of data

-   summarise() summarises data

-   arrange()

-   mutate()

```{r}
df <- read_csv('https://bryantstats.github.io/math421/data/all-states-history.csv')
```

na.rm = TRUE does calculation with nas

```{r}
df %>% 
  mutate(wd2 = weekdays(date)) %>% 
  filter(state == "RI") %>% 
  group_by(wd2) %>% 
  summarise(median_death_RI = median(death, na.rm=TRUE))


  
```

```{r}
df %>% 
  mutate(wd2 = weekdays(date)) %>% 
  filter(state == "RI") %>% 
  group_by(wd2) %>% 
  summarise(median_death_RI = median(death, na.rm=TRUE)) %>% 
  arrange(-median_death_RI)

```

```{r}
df$wd = weekdays(df$date)

#filter out RI

df1 = df[df$state=="RI",]

#calculate by groups

by(data = df$death, INDICES = df$wd, median)
```

```{r}
df2 = df1[df1$wd == "Sunday",]
mean(df2$death)
mean(df2$death, na.rm = TRUE)
median(df2$death, na.rm = TRUE)
```
