# Data Wrangling with Base R

## Data Frame

-   table of data

-   each column contains values of one variable

-   each row contains one set of values from each column

## creating a data frame

-   use the data.frame() function

```{r}
df = data.frame(Name = c("A", "B"), Age = c(20,30))
df
```

```{r}
df = data.frame(Name = c('Rick','Dan','Jessica','Julia','Sophia'),
                Salary = c(623.30,515.20,611.00,729.00,843.25),
                Start_Date = c('2012-01-01',
                                       '2013-09-23',
                                       '2014-11-15',
                                       '2014-05-11',
                                       '2015-03-27'),
                Gender = c('Male', 'Male', 'Female','Female','Female' ))
df
```

## Functions

**import:** read.csv

```{r}
df = read.csv("WHO-COVID-19-global-data.csv")
```

**overview:** str

```{r}
str(df)
```

-   gives type of data dealing with

-   quality of data (missing values)

**correct data type:**

-   as.factor

-   as.numeric

-   as.Date

-   as.character

**check missing data**

-   sum(is.na())

-   checks number of missing values

```{r}
sum(is.na(df))
```

-   colSums((is.na))

-   values in each column

```{r}
colSums(is.na(df))
```

## changing the name of a variable

```{r}
names(df)
```

```{r}
names(df)[3] = "country_name"
```

## change the type of variable

check type

```{r}
class(df$Date_reported)
```

change type

```{r}
df$Date_reported = as.Date(df$Date_reported)
```

## Simple Data Exploration

```{r}
summary(df)
```

### Statistics of continuous / date variable

```{r}
mean(df$Cumulative_deaths, na.rm = TRUE)
```

```{r}
max(df$Cumulative_deaths, na.rm = TRUE)
```

```{r}
summary(df$Date_reported)
```

### frequency of categorical variable

```{r}
table(df$WHO_region)
```

```{r}
# % of the variable for each type
prop.table(table(df$WHO_region))
```

### correlation of two continuous variables

```{r}
cor(df$Cumulative_cases, df$Cumulative_deaths)
```

### calculate a continuous by a categorical

```{r}
by(df$New_deaths, df$WHO_region, FUN = mean)
```

## create new variables

*it is common to create new variables to analyze the data better*

### 

### binning continuous variable to a categorical variable

```{r}
df$Cases2 <- ifelse(df$Cases==0, 'No_new_cases', 'Has_new_cases')
```

```{r}
str(df)
```

can we create a new variable with three categories using ifelse function?

```{r}
df$Cases3 = case_when(df$Cases < 22 ~ "Low",
                      df$Cases < 486 ~ "Medium",
                      TRUE ~ "High")
```

```{r}
table(df$Cases3)
```

```{r}
prop.table(table(df$Cases3))
```

## **binning continuous variable**

creates different levels

```{r}
df$Death2 <- case_when(df$Death < 2 ~ 'low_death',
                       df$Death < 5 ~ 'mid_death',
                       TRUE ~ 'high_death')

table(df$Death2)
```

## Subsetting by Columns

```{r}
df1 <- df[,c(1,3)]
head(df1)
```

only contains days with cases greater than 1000

```{r}
df_high_cases =df[df$Cases > 1000, ]
```

```{r}
str(df_high_cases)
```
