---
title: "Midterm"
author: "Aileen Charron"
date: "2024-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Wrangling

## 1.1 Download the data file

## 1.2 Read the dataset

```{r}
#import libraries
library(haven)
library(tidyverse)
library(VIM)
library(dplyr)
library(lubridate)
library(caret)
library(rpart)
library(rattle)
library(mlbench)
library(gridExtra)
library(ggthemes)
library(gganimate)
library(gifski)
library(knitr)
library(gapminder)
library(VIM)
```

```{r, eval=FALSE}
#read data
df = read_sas("C:\\Users\\student\\Documents\\M461-Statistical-Analysis-with-R\\hdd0318cy.sas7bdat")
```

## 1.3 Filter the data to have only patients of the year 2018 (yod = 18)

```{r, eval=FALSE}
#filter year to 2018
df <- df %>% 
  filter(yod == 18)
```

## 1.4 Select to work with only following variables

```{r, eval=FALSE}
#select variables
df <- df %>% 
  select("yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day")


```

## 1.5 What variables that have missing values?

```{r}

df <- read.csv("C:\\Users\\student\\Documents\\M461-Statistical-Analysis-with-R\\midtermdata.csv")
#determining missing values
colSums(is.na(df))
```

-   payfix

-   preopday

-   obs_hour

-   nicu_day

## 1.6 Remove all variables with missing values

```{r}
#removing variables with missing variables
df = df %>% select(-"payfix", -"preopday", -"obs_hour", -"nicu_day")
```

## 1.7 Which variable recording the month of admission? Which variable recording the month of discharge?

-   moa: month of admission

-   mod: month of discharge

```{r}
#creating month of admission variable with month names
df$moa_name <- case_when(
  df$moa  == 1 ~ "Jan",
  df$moa == 2 ~ "Feb",
  df$moa == 3 ~ "Mar",
  df$moa == 4 ~ "Apr",
  df$moa == 5 ~ "May",
  df$moa == 6 ~ "Jun",
  df$moa == 7 ~ "Jul",
  df$moa == 8 ~ "Aug",
  df$moa == 9 ~ "Sep",
  df$moa == 10 ~ "Oct",
  df$moa == 11~ "Nov",
  df$moa == 12 ~ "Dec")

df <- df %>% 
  mutate(moa_name = factor(moa_name, levels = month.abb))
```

```{r}
#creating month of discharge variable with month names
df$mod_name <- case_when(
  df$mod  == 1 ~ "Jan",
  df$mod == 2 ~ "Feb",
  df$mod == 3 ~ "Mar",
  df$mod == 4 ~ "Apr",
  df$mod == 5 ~ "May",
  df$mod == 6 ~ "Jun",
  df$mod == 7 ~ "Jul",
  df$mod == 8 ~ "Aug",
  df$mod == 9 ~ "Sep",
  df$mod == 10 ~ "Oct",
  df$mod == 11~ "Nov",
  df$mod == 12 ~ "Dec")

df <- df %>% 
  mutate(mod_name = factor(mod_name, levels = month.abb))
```

```{r}
#creating provider name varialbe
df$provider_name <- case_when(
  df$provider == 7201 ~ "Newport",
  df$provider == 7202 ~ "St. Joseph",
  df$provider == 7203 ~ "Memorial",
  df$provider == 7204 ~ "Miriam",
  df$provider == 7205 ~ "Rhode Island Hospital",
  df$provider == 7206 ~ "Roger Williams",
  df$provider == 7209 ~ "South County",
  df$provider == 7210 ~ "Kent County",
  df$provider == 7211 ~ "Westerly",
  df$provider == 7212 ~ "Rehab of RI",
  df$provider == 7213 ~ "Landmark Medical Center",
  df$provider == 7214 ~ "Women and infants",
  df$provider == 7215 ~ "Bradley",
  df$provider == 7216 ~ "Butler"

)
```

```{r}
#creating age group variable

df$age_group <- case_when(
  df$age <= 10 ~ "0-10",
  df$age <= 20 ~ "11-20",
  df$age <= 30 ~ "21-30",
  df$age <= 40 ~ "31-40",
  df$age <=50 ~"41-50",
  df$age <= 60 ~ "51 - 60",
  df$age <= 70 ~ "61 - 70",
  df$age <= 80 ~ "71 - 80",
  df$age <= 90 ~ "81 - 90",
  df$age <= 100 ~ "91 - 100",
  df$age > 100 ~ "100+"
)
```

```{r}
#removing observations where sex = 9
df <- df %>% filter(sex != 9)
```

```{r}
#creating a sex name variable

df$sex_name <- case_when(
  df$sex == 1 ~ "Male",
  df$sex == 2 ~ "Female")
```

```{r}
#creating variable that indicates 1 if in providence, 0 if not
df <- df %>%
  mutate(Providence = case_when(
    provider %in% c(7205, 7206, 7209, 7213, 7216) ~ 1,
    TRUE ~ 0
  ))
```

```{r}
#making Providence as factor
df <- df %>% 
  mutate(Providence = as.factor(Providence))
```

## 1.8 Which month admitted the most number of patients? Which month admitted the most number of male patients?

```{r}
#month with most admitted patients
df %>% group_by(moa_name) %>% 
  summarize(total_admissions = n()) %>% 
  arrange(desc(total_admissions))
```

October admitted the most patients

```{r}
#month with most admitted male patientw
df %>%  filter(sex == 1) %>% 
  group_by(moa_name) %>% 
  summarize(total_admissions = n()) %>% 
  arrange(desc(total_admissions))

```

October admitted the most male patients

## 1.9 Which month has the most number of teenage female patients?

```{r}
#month with most teenage female patients
df %>% filter(sex == 2) %>% 
filter(age >= 13) %>% 
  filter(age <=19) %>% 
  group_by(moa_name) %>% 
  summarize(total_admissions = n()) %>% 
  arrange(desc(total_admissions))
```

March admitted the most teenage female patients

## 1.10 which provider has the most number of female patients in October?

```{r}
df %>% filter(sex == 2) %>% 
  group_by(provider_name) %>% 
  summarize(total_admissions = n()) %>% 
  arrange(desc(total_admissions))
```

Rhode Island hospital had the most number of female patients in October

## 1.11 are female patients older than male patients, on average?

```{r}
df %>% group_by(sex_name) %>% 
  summarize(average_age = mean(age)) %>% 
  arrange(average_age)
```

On average, males are older than females

## 1.12 calculate the average age of patients by month. which month has the oldest patients on average age?

```{r}
df %>% group_by(moa_name) %>% 
  summarize(average_age = mean(age)) %>% 
  arrange(desc(average_age))
```

January has the oldest patients on average

## 1.13 what is the name of the provider that has the highest total charge?

```{r}
df %>% group_by(provider_name) %>% 
  summarize(average_total_charge = mean(tot)) %>% 
  arrange(desc(average_total_charge))
  
```

Bradley has the highest average total charge

## 1.14 what is the name of the provider that has the least total charge for teenage male on average

```{r}
df %>% 
  filter(sex == 1) %>% 
  filter(age >= 13 | age <=19) %>% 
  group_by(provider_name) %>% 
  summarize(average_total_charge = mean(tot)) %>% 
  arrange(average_total_charge)
```

Butler has the least total charge for teenage males on average

## 1.15 create a season (spring, summer, fall, winter) variable. calculate the length of the stays by season. which season has the longest of stays on average?

```{r}
df$season <- case_when(
  df$moa %in% c(7,8,9) ~ "summer",
  df$moa %in% c(10,11,12) ~ "fall",
  df$moa %in% c(1,2,3) ~ "winter",
  df$moa %in% c(4,5,6) ~ "spring")
```

```{r}
df %>% 
  group_by(season) %>% 
  summarize(average_los = mean(los)) %>% 
  arrange(desc(average_los))
```

fall has the longest average length of stay on average

## 1.16 on average, how much do a 20 year old male get charged for staying 1 day in the Fall season?

```{r}
df %>% 
  filter(age == 20) %>% 
  filter(sex == 1) %>% 
  filter(los == 1) %>% 
  filter(season == "fall") %>% 
  summarize(average_tot = mean(tot)) 
```

on average, a 20 year old male gets charged \$16,511 for staying 1 day in the fall

## 1.17 write a paragraph to summarize the section and give your comments on the results. you could do some other calculations to support your points

Rhode Island hospital had the most number of female patients in October. On average, males are older than females. January has the oldest patients on average. Bradley has the highest average total charge. Butler has the least total charge for teenage males on average. fall has the longest average length of stay on average. on average, a 20 year old male gets charged \$16,511 for staying 1 day in the fall

The data definitely shows variation within the data when subsetted into different subcategories.

# 2. Data Visualization

## 2.1 provide at least 10 meaningful plots. comment on the plots. all plots should have title, caption, appropriate labels on x and y- axis

```{r}
#1
# looking at total admissions by month by gender
graph1 <- df %>% ggplot + geom_bar(mapping = aes(x = moa_name, fill = sex_name), position = 'dodge') + 
  labs(
    title = "Admissions by Month",
    x = "month",
    y = "admissions",
    caption = "In all months, there are more female admissions than male admissions"
  ) + theme_economist()

ggsave(filename = 'graph1.png', plot = graph1)

graph1

```

```{r}
#2
# looking at total charges by Providence or not
graph2 <- df %>% group_by(moa_name, Providence) %>% 
  summarize(average_tot = mean(tot)) %>%
  ggplot + 
  geom_col(mapping = aes(x = moa_name, y = average_tot, fill = Providence), position = 'dodge') + 
  labs(
    title = "Average Total Charge By Month",
    x = "month",
    y = "average total charge",
    caption = "In all months, the average total charge is higher for Hospitals in Providence than those not in Providence"
  ) + theme_economist()

ggsave(filename = 'graph2.png', plot = graph2)

graph2

```

```{r}
#3
# looking at age vs LOS by gender
graph3 <- df %>% group_by(age, sex_name,los) %>% 
  summarize(average_LOS = mean(los)) %>% 
  ggplot(aes(x = age, y = average_LOS))+
  geom_point() +
  labs(
    title = "Age vs Average LOS by Gender",
    x = "Age",
    y = "Average Length of Stay",
    caption = "Length of stay does not appear to vary by age nor gender with the exception of newborns") +
  facet_wrap(~sex_name)+
  theme_economist()

ggsave(filename = 'graph3.png', plot = graph3)

graph3

```

```{r}
# 4
#looking at age vs avg total charge by gender
graph4 <-df %>% group_by(age, sex_name) %>% 
  summarize(average_tot = mean(tot)) %>% 
  ggplot(aes(x = age, y = average_tot, color = sex_name))+
  geom_point() +
  labs(
    title = "Age vs Average total charge by Gender",
    x = "Age",
    y = "Average total charge",
    caption = "appears to have a dip aorund age 30 for both male and females, females tend to be slightly lower charges") +
  theme_economist()

ggsave(filename = 'graph4.png', plot = graph4)

graph4

```

```{r}
# 5 looking at total admissions by provider by gender
graph5 <- df %>% group_by(provider_name, sex_name) %>% 
  ggplot + 
  geom_bar(mapping = aes(x = provider_name, fill = sex_name), position = 'dodge') + 
  labs(
    title = "Total Admissions by Provider by Gender",
    x = "Provider",
    y = "Admissions",
    caption = "Rhode Island Hospital Has the Most Admissions"
  ) + theme_economist() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8)
  )
ggsave(filename = 'graph5.png', plot = graph5)

graph5



```

```{r}
# 6 
#looking at average total charge by provider by gender
graph6 <-df %>% group_by(provider_name, sex_name) %>% 
  summarize(average_total_charge = mean(tot)) %>% 
  ggplot + 
  geom_col(mapping = aes(x = provider_name, y = average_total_charge, fill = sex_name), position = "dodge") + 
  labs(
    title = "average total charge by Provider by Gender",
    x = "Provider",
    y = "average total charge",
    caption = "Bradley has highest average charge") + theme_economist() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8)) +
      scale_y_continuous(labels = scales::comma)

graph6

ggsave(filename = 'graph6.png', plot = graph6)

```

```{r}
#7 

#comparing average total charges of different ER modes by month

graph7 <- df %>% 
  group_by(er_mode, moa_name) %>% 
  summarize(average_total_charge = mean(tot)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = moa_name, 
                           y = average_total_charge,
                           color = er_mode)) + 
  labs(
    title = "average total charge ER Mode",
    x = "Month",
    y = "average total charge",
    caption = "ER Modes 2 & 3 have highest average charges each month (helicopter & law enforcement agency)") + theme_economist() +
      scale_y_continuous(labels = scales::comma)

graph7

ggsave(filename = 'graph7.png', plot = graph7)


```

```{r}
#8

# comparing frequency of ER Modes by month
graph8 <- df %>% 
  filter(er_mode %in% c(2,3)) %>% 
  group_by(provider_name, er_mode) %>% 
  summarize(frequency = n(), .groups = "drop") %>% 
  ggplot(aes(x = provider_name, y = frequency, fill = er_mode)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Frequency of ER Modes 2 & 3 by Provider",
       x = "provider",
       y = "frequency",
       fill = "ER mode",
       caption = "only Lankmark & RI Hospital had helicopter ER modes in 2018") +
      theme_economist() +
    theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8)) +
      scale_y_continuous(labels = scales::comma)

ggsave(filename = 'graph8.png', plot = graph8)


graph8
```

```{r}
# 9
 # patient frequency by state
graph9 <-df %>% filter(pt_state %in% c("RI", "MA", "CT", "NH", "ME", "VT")) %>% 
  group_by(pt_state, provider_name) %>% 
  summarize(frequency = n(), .groups = "drop") %>% 
  ggplot + 
  geom_col(mapping = aes(x = provider_name, y = frequency, fill = pt_state), position = "dodge") + 
  labs(
    title = "Frequency of Patients from New England States",
    x = "Provider",
    y = "frequency",
    caption = "Most Patients are from RI or MA") + theme_economist() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8)) +
      scale_y_continuous(labels = scales::comma)

ggsave(filename = 'graph9.png', plot = graph9)


```

```{r}
 #dist of patient birthweights
graph10<-df %>%
  filter(b_wt != "000") %>% 
  ggplot(aes(x = as.numeric(b_wt))) +  
  geom_histogram(binwidth = 10, fill = "pink", color = "black") + 
  labs(
    title = "Birth Weights Distribution",
    x = "weight in grams at birth",
    y = "Frequency",
    caption = "birth weight has a normal distribution"
  ) +
  theme_economist()

graph10
ggsave(filename = 'graph10.png', plot = graph10)



```

## 2.2 make an animation plot

```{r}
d1 <- df %>% 
  group_by(moa_name, sex_name) %>% 
  summarize(frequency = n(), .groups = "drop")
d2 <- d1 %>% 
  group_by(moa_name) %>% 
  mutate(rank = rank(-frequency))

a1 <- d2 %>% 
  ggplot(aes(x = rank, y = frequency, group = sex_name, fill = sex_name, label = sex_name)) +
  geom_col() +
  geom_text(aes(
    y = frequency,
    label = sex_name),
    hjust = 1.4) + 
  coord_flip(clip = "off", expand = FALSE) +
  scale_x_reverse() +
  labs(title = 'Month {closest_state}',
       x = '',
       y = 'total admissions',
       fill = 'Sex') +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  transition_states(moa_name) + 
  ease_aes("cubic-in-out")

animate(a1, nframes = 400)
```

```{r}
anim_save("animatedgif.gif", a1)
```

## 2.3 write a paragraph to summarize the section and give your comments on the results

Like section 1, section 2 showed that there are key trends that emerge when the data is filtered and presented in different ways. visualization across providers, months, age, and gender highlighted these key trends. This can be useful for hospitals and doctors as they can come to anticipate these trends within subsets of their patient populations.

# 3. predictive models

## 3.0 create the target variable

-   taking value of low if total charge of a patient (tot) is smaller than the median of total charge, high otherwise

```{r}
median_tot = median(df$tot)

df$target <- case_when(
  df$tot <= median_tot ~ 'low',
  df$tot > median_tot ~ 'high'
)
```

## 3.1 split data

```{r}
# filter out rows were raceethn == '' or admtype == ''

df <- df %>% filter(
  raceethn != "")

df <- df %>% filter(
  admtype != ""
)

df_inputs <- df %>% select("age", "sex", "raceethn", "provider", "moa", "mod", "admtype", "campus", "los", "target")

# categorical = factor, numeric = numeric

df_inputs <- df_inputs %>% 
  mutate(target = as.factor(target),
         age = as.numeric(age),
         sex = as.factor(sex),
         raceethn = as.factor(raceethn),
         moa = as.factor(moa),
         mod = as.factor(mod),
         admtype = as.factor(admtype),
         campus = as.factor(campus),
         los = as.numeric(los))

# set training split
set.seed(2020)

invisible(splitIndex <- createDataPartition(df_inputs$target, p =0.1,
                                 list = FALSE))

df_train <- df_inputs[splitIndex, ]
df_test <- df_inputs[-splitIndex, ]
```

## 3.2

-   train a decision tree using rpart

```{r}
dt <- rpart(target ~., data = df_train)
```

-   plot the decision tree

```{r}
fancyRpartPlot(dt)
```

-   plot variable importance ranked by the tree

```{r}
barplot(dt$variable.importance)
```

## 3.3 cross validation 5 folds

```{r}
# k control is cross validation of 5 folds
trControl = trainControl(method = "cv",
                         number = 5)


```

```{r}
# rpart with train control
dt2 <- train(target ~., data = df_train,
             method = "rpart",
             trControl = trControl)
```

-   training & tune at least 2 different models (two different values for method= in the train function of caret)

```{r}
# k nearest neighbors
knn <- train(target ~., data = df_train,
              method = "knn",
              trControl = trControl)
```

```{r}
ranger <- train(target ~., data = df_train,
            method = "ranger",
            trControl = trControl)
```

-   plot the hyper parameter tuning plots for each model

```{r}
plot(dt2)
plot(ranger)
plot(knn)
```

## 3.4

-   plot the comparison of the models in 3

```{r}
results <- resamples(list ('decision tree' = dt2,
                           'k nearest neighbors' = knn,
                           'ranger' = ranger
                ))

bwplot(results)


```

## 3.5

-   what is your final selection for the model? test the accuracy of your final model on the test data

ranger

```{r}
pred <- predict(ranger, df_test)

cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "high")

cm$overall[1]
```

## 3.6

-   create another target variable (binary), decide the input variables and redo 1 to 5

new target is if los is greater than median or not

```{r}
median_los = median(df$los)

df$target2 <- case_when(
  df$los <= median_los ~ 'low',
  df$los > median_los ~ 'high'
)
```

### 3.6.1

-   use filter function to filter out rows were reaceethn=='' or admtype == ''. Make sure all the categorical variables are factor, numeric variables are numeric. set training: testing split = 10:90

```{r}
df_inputs2 <- df %>% select("age", "sex", "raceethn", "provider", "moa", "mod", "admtype", "campus", "tot", "target2", "season")

# categorical = factor, numeric = numeric

df_inputs2 <- df_inputs2 %>% 
  mutate(target2 = as.factor(target2),
         age = as.numeric(age),
         sex = as.factor(sex),
         prover = as.factor(provider),
         raceethn = as.factor(raceethn),
         moa = as.factor(moa),
         mod = as.factor(mod),
         admtype = as.factor(admtype),
         campus = as.factor(campus),
         season = as.factor(season))
# set training split
set.seed(2020)

invisible(splitIndex2 <- createDataPartition(df_inputs2$target2, p =0.1,
                                 list = FALSE))

df_train2 <- df_inputs2[splitIndex2, ]
df_test2 <- df_inputs2[-splitIndex2, ]
```

### 3.6.2

-   train a decision tree using rpart

```{r}
dt3 <- rpart(target2 ~., data = df_train2)
```

-   plot the decision tree

```{r}
fancyRpartPlot(dt3)
```

-   plot variable importance ranked by the tree

```{r}
barplot(dt3$variable.importance)
```

### 3.6.3

-   using caret for this question, set raining control to be: use cross validation of 5 folds across all models

```{r}
# k control is cross validation of 5 folds
trControl = trainControl(method = "cv",
                         number = 5)
```

```{r}
# rpart with train control
dt4 <- train(target2 ~., data = df_train2,
             method = "rpart",
             trControl = trControl)
```

-   training & tune at least 2 different models (two different values for method= in the train function of caret)

```{r}
# random forest
invisible(nnet <- train(target2 ~., data = df_train2,
             method = "nnet",
             trControl = trControl))

```

```{r}
# k nearest neighbors
ranger <- train(target2 ~., data = df_train2,
              method = "ranger",
              trControl = trControl)
```

-   plot the hyper parameter tuning plots for each model

```{r}
plot(dt4)
plot(nnet)
plot(ranger)
```

### 3.6.4

-   plot the comparison of the models in 3

```{r}
results <- resamples(list ('decision tree' = dt4,
                           'neural network' = nnet,
                           'ranger' = ranger))
bwplot(results)
```

### 3.6.5

```{r}
pred2 <- predict(ranger, df_test2)

cm2 <- confusionMatrix(data = pred2, reference = df_test2$target2, positive = "high")

cm2$overall[1]
```

## 3.7

-   write a paragraph to summarize the section and give your comments on the results

length of stay and total charge are very correlated variables and are the most important predictors of each other. the location/provider and admission type also are important predictors of length of stay. as for total charge, admission type and provider are important predictors as well. ranger proved to be the most successful method of predicting both of my chosen target variables.
